---
title: "Modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(GGally)
library(knitr)
library(gridExtra)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `movies`. Delete this note when before you submit 
your work. 

```{r load-data}
load("movies.Rdata")
```



* * *

## Part 1: Data
### <span style="color: brown">What is this data about?  </span>  

- The dataset contains a list of randomly sampled movies from the IMDB and Rotten Tomatoe movie databases.  

- It contains information about both audience and critic scores about how much they like/dislike movies. As well as other variables associated with the movie.  

- The dataset contains 651 rows and 32 variables.  

- The codebook can be viewed from this link: [Code Book](https://d18ky98rnyall9.cloudfront.net/_73393031e98b997cf2445132f89606a1_movies_codebook.html?Expires=1586908800&Signature=IZy8756SrMG5MDxovj4cf9XfLTeDQzOuNK250lyE~TgJ1sfjmrUJ17K2jc3MdX~NsFgNJogFeS4K-7iKNRlFmQzdT~zaW1NplyDa9vc8uUCvBtq1RXHKparTVJxQPfAItI0ol4WCvW-wNE1qiuT37XJ6~5ZYqd75nJEvoZMtsnI_&Key-Pair-Id=APKAJLTNE6QMUY6HBC5A)

#### <span style="color: red">Sampling: </span> 

- **Random Sampling** was used in the selection of the movies from the IMDB and Rotten Tomatoe databases.

#### <span style="color: red">Generalizability and Association vs Causation</span>

<span style="color: Green">**Generalizability** :</span>

- The survey uses data from the Rotten Tomatoes and IMDB databases.

- These are some of the biggest databases in the world, that has data on a huge repository of movies from around the world.

- However, on closer inspection. The data about the movies in the sample contain only English language movies.  

### <span style="color: Orange">**Answer : **</span>
- So we can safely say that the sample data is generalizable **only** to English language movies from the IMDB and Rotten Tomato database.  

<span style="color: Green">**Causality** :</span>

- The data is **Associative** at the moment.

- However, controlled random assignment of new movies to any model generated, can be used to establish **causal** relationships between **response** and **explanatory** variables.

### <span style="color: Orange">**Answer : **</span>

- The data is **Associative** at the moment.  

- **However** ,there is a possibilty of drawing conclusions on **causality** only after conducting many simulations on generated models to figure out there is strong evidence of **causality** between the **response** and **explanatory** variables used to create the model.



* * *

## Part 2: Research question
#### <span style="color: Brown">**What are the best predictor variables to create an accurate model to predict an IMDB score ?** </span> 

- Real life modelling involves,looking at lots of variables from different sources and piecing them together to make a robust model that can predict an acurate outcome at a specified confidence interval. 

- I am interested in creating a predictor with a 95% confidence interval to predict the IMDB score of a movie, based on various predictor variables.  

- I am planning to create the most efficient parsimonious model, which can explain most of the variability in the IMDB ratings.  

- <span style="color: Green">**In order to achieve this , I will be using a forward selection approach, with emphasis on getting the best adjusted R-squared value.**</span>


* * *

## Part 3: Exploratory data analysis
### <span style="color: Orange">**Step1 : **</span>  

#### <span style="color: darkgreen">**We begin by having an overview of the model:** </span> 

```{r}
str(movies)
```
### <span style="color: Brown">**Result: **</span> 

- We see there are **10 numeric variables** in the dataset:  

- title,actor1-actor5 , imdb_url, rt_url are **9 character variables**.  

- there are **12 factor variables** in the dataset.

- there is **1 integer variable** imdb_num_votes. 


### <span style="color: Orange">**Step2 : **</span>  

#### <span style="color: darkgreen">**Let us now plot a ggpairs plot of all the numeric variables in our dataset.** </span> 


```{r, fig.width = 12, fig.height= 12, message=FALSE, warning=FALSE, cache=TRUE}
m1 <- movies%>%select(c(runtime,thtr_rel_year,thtr_rel_month,thtr_rel_day,dvd_rel_year,dvd_rel_month,dvd_rel_day,imdb_rating,imdb_num_votes,critics_score,audience_score))

ggpairs(data = m1)
```

#### <span style="color: red">**ANOVA** </span>
```{r}
AnoV <- aov(imdb_rating~imdb_num_votes+critics_score+audience_score+runtime+thtr_rel_year+thtr_rel_month+thtr_rel_day+dvd_rel_year+dvd_rel_month+dvd_rel_day, data = m1)
summary(AnoV)
```
```{r, include=FALSE}
lm2_1 <- lm(imdb_rating~imdb_num_votes+critics_score+audience_score+runtime+thtr_rel_year+thtr_rel_month+thtr_rel_day+dvd_rel_year+dvd_rel_month+dvd_rel_day, data = m1)
#summary(lm2_1)$r.squared
#summary(lm2_1)$adj.r.squared
```

<span style="color: red">**Multiple R-squared:** </span> **`r summary(lm2_1)$r.squared`** 

<span style="color: red">**Adjusted R-squared:** </span> **`r summary(lm2_1)$adj.r.squared`**

- We observe from ANOVA: critics_score, audience_score, imdb_num_votes, runtime. Contribute maximum to the variability, due to ther high values.

- We will rebuild a model with only these 4, to see what the R values look like.

```{r}
AnoV <- aov(imdb_rating~imdb_num_votes+critics_score+audience_score+runtime, data = m1)
summary(AnoV)
```

```{r, include=FALSE}
lm2_1 <- lm(imdb_rating~imdb_num_votes+critics_score+audience_score+runtime, data = m1)
summary(lm2_1)
```


<span style="color: red">**Multiple R-squared:** </span> **`r summary(lm2_1)$r.squared`** 

<span style="color: red">**Adjusted R-squared:** </span> **`r summary(lm2_1)$adj.r.squared`**



### <span style="color: Brown">**Result:  Why did we pick only these 4 numeric variables **</span> 

- We see that this parsimonious model has a better overall adjusted R-squared

- Therefore, these 4 numeric variables will be analyzed in depth, to build out model.

- Also these 4 numeric variables have the highest correlation coefficients, which indicated they have a more linear relationship. The other numeric variables can be dropped from consideration since they do not have strong linear relationsips.

- I am **NOT** taking the critics_score and audience_score collinearity. Because there has been times when critics and audiences have diverging scores. So even though the correlation seems high, based on experience. I am not going to use collineartiy to drop one of them. Besides there must be some really important information that seems to superceded the collinearity condition, as the adjusted_R squared value has actually gone up substantially.




### <span style="color: Orange">**Step3 : **</span>  

#### <span style="color: darkgreen">plot **IMDB rating** against the **11 factor variables(excluding studio)** to see if we can learn something: </span> 


```{r, fig.width = 12, fig.height= 12, message=FALSE, warning=FALSE,cache=TRUE}
m2 <- movies%>%select(c(imdb_rating,title_type,genre,mpaa_rating,critics_rating,audience_rating,best_pic_nom,best_pic_win,best_actor_win,best_actress_win,best_dir_win,top200_box))

ggpairs(data = m2)

```

### <span style="color: Brown">**Result: **</span> 

- Looking at the first row, we can see that most categorical variables exhibit a high degree of variance against the imdb_rating. This is something interesting that needs to be noted.  the 3 that seem to have the least variance are: best_actor_win, best_actress_win, best_director_win.  

* * *

#### <span style="color: red">ANOVA and adjusted-R-Squared: </span> <span style="color: darkgreen"> **imdb_rating vs title_type** </span>
```{r}
AnoV <- aov(imdb_rating~title_type, data = m2)
summary(AnoV)
```
```{r, include=FALSE}
lm2_1 <- lm(imdb_rating~title_type, data = m2)
summary(lm2_1)
```

### <span style="color: Brown">**Result:**</span> 

<span style="color: red">**Multiple R-squared:** </span> **`r summary(lm2_1)$r.squared`** 

<span style="color: red">**Adjusted R-squared:** </span> **`r summary(lm2_1)$adj.r.squared`**

<span style="color: brown">**Note** adjusted R-squared has not dropped significantly, this seems like a good candidate for our model</span>  

* * *

#### <span style="color: red">ANOVA and adjusted-R-Squared: </span> <span style="color: darkgreen"> **imdb_rating vs genre** </span>
```{r}
AnoV <- aov(imdb_rating~genre, data = m2)
summary(AnoV)
```
```{r, include=FALSE}
lm2_1 <- lm(imdb_rating~genre, data = m2)
summary(lm2_1)
```

### <span style="color: Brown">**Result:**</span> 

<span style="color: red">**Multiple R-squared:** </span> **`r summary(lm2_1)$r.squared`** 

<span style="color: red">**Adjusted R-squared:** </span> **`r summary(lm2_1)$adj.r.squared`**

<span style="color: brown">**Note** adjusted R-squared has not dropped significantly, this seems like a good candidate for our model</span>  

* * *

#### <span style="color: red">ANOVA and adjusted-R-Squared: </span> <span style="color: darkgreen"> **imdb_rating vs mpaa_rating** </span>
```{r}
AnoV <- aov(imdb_rating~mpaa_rating, data = m2)
summary(AnoV)
```

```{r, include=FALSE}
lm2_1 <- lm(imdb_rating~mpaa_rating, data = m2)
summary(lm2_1)
```


### <span style="color: Brown">**Result:**</span> 

<span style="color: red">**Multiple R-squared:** </span> **`r summary(lm2_1)$r.squared`** 

<span style="color: red">**Adjusted R-squared:** </span> **`r summary(lm2_1)$adj.r.squared`**

<span style="color: brown">**Note** mpaa_rating seems to have a verly low R-squared and an even lower adjusted R-squared. We may ignore this while buiilding the model.</span> 

* * *

#### <span style="color: red">ANOVA and adjusted-R-Squared: </span> <span style="color: darkgreen"> **imdb_rating vs critics_rating** </span>
```{r}
AnoV <- aov(imdb_rating~critics_rating, data = m2)
summary(AnoV)
```

```{r, include=FALSE}
lm2_1 <- lm(imdb_rating~critics_rating, data = m2)
summary(lm2_1)
```

### <span style="color: Brown">**Result:**</span> 

<span style="color: red">**Multiple R-squared:** </span> **`r summary(lm2_1)$r.squared`** 

<span style="color: red">**Adjusted R-squared:** </span> **`r summary(lm2_1)$adj.r.squared`**

<span style="color: brown">**Note** adjusted R-squared has not dropped significantly, this seems like a good candidate for our model</span>  

* * *

#### <span style="color: red">ANOVA and adjusted-R-Squared: </span> <span style="color: darkgreen"> **imdb_rating vs audience_rating** </span>

```{r}
AnoV <- aov(imdb_rating~audience_rating, data = m2)
summary(AnoV)
```

```{r, include=FALSE}
lm2_1 <- lm(imdb_rating~audience_rating, data = m2)
summary(lm2_1)
```


### <span style="color: Brown">**Result:**</span> 

<span style="color: red">**Multiple R-squared:** </span> **`r summary(lm2_1)$r.squared`** 

<span style="color: red">**Adjusted R-squared:** </span> **`r summary(lm2_1)$adj.r.squared`**

<span style="color: brown">**Note** adjusted R-squared has not dropped significantly, this seems like a good candidate for our model</span>  

* * *

#### <span style="color: red">ANOVA and adjusted-R-Squared: </span> <span style="color: darkgreen"> **imdb_rating vs best_pic_nom** </span>
```{r}
AnoV <- aov(imdb_rating~best_pic_nom, data = m2)
summary(AnoV)

lm2_1 <- lm(imdb_rating~best_pic_nom, data = m2)
summary(lm2_1)
```

<span style="color: red">**Note** adjusted R-squared has not dropped significantly, this seems like a good candidate for our model</span> 

* * *

#### <span style="color: red">ANOVA and adjusted-R-Squared: </span> <span style="color: darkgreen"> **imdb_rating vs best_pic_win** </span>
```{r}
AnoV <- aov(imdb_rating~best_pic_win, data = m2)
summary(AnoV)

lm2_1 <- lm(imdb_rating~best_pic_win, data = m2)
summary(lm2_1)
```

<span style="color: red">**Note** This seems like a decent candidate, but might be dropped as adjusted R-squared is quite small</span> 

* * *

#### <span style="color: red">ANOVA and adjusted-R-Squared: </span> <span style="color: darkgreen"> **imdb_rating vs best_actor_win** </span>
```{r}
AnoV <- aov(imdb_rating~best_actor_win, data = m2)
summary(AnoV)

lm2_1 <- lm(imdb_rating~best_actor_win, data = m2)
summary(lm2_1)
```

<span style="color: red">**Note** adjusted R-squared is too small to be of significance, this looks to be definitely dropped. This corresponds to the lack of noticable variance in the box-plots.</span> 

* * *

#### <span style="color: red">ANOVA and adjusted-R-Squared: </span> <span style="color: darkgreen"> **imdb_rating vs best_actress_win** </span>
```{r}
AnoV <- aov(imdb_rating~best_actress_win, data = m2)
summary(AnoV)

lm2_1 <- lm(imdb_rating~best_actress_win, data = m2)
summary(lm2_1)
```

<span style="color: red">**Note** adjusted R-squared is too small to be of any value, we will ignore this while building the model</span> 

* * *

#### <span style="color: red">ANOVA and adjusted-R-Squared: </span> <span style="color: darkgreen"> **imdb_rating vs best_dir_win** </span>
```{r}
AnoV <- aov(imdb_rating~best_dir_win, data = m2)
summary(AnoV)

lm2_1 <- lm(imdb_rating~best_dir_win, data = m2)
summary(lm2_1)
```

<span style="color: red">**Note** adjusted R-squared is really small, we will be dropping this while building the model</span> 

* * *

#### <span style="color: red">ANOVA and adjusted-R-Squared: </span> <span style="color: darkgreen"> **imdb_rating vs top200_box** </span>
```{r}
AnoV <- aov(imdb_rating~top200_box, data = m2)
summary(AnoV)

lm2_1 <- lm(imdb_rating~top200_box, data = m2)
summary(lm2_1)
```

<span style="color: red">**Note** adjusted R-squared os really small. We won't be including this in our model</span> 

<span style="color: red"></span>
 
* * *

### <span style="color: Orange">**Step4 : **</span>  

#### <span style="color: darkgreen">**Now we can look at the relationship between IMDB_ratings and studios next:** </span> 

```{r, fig.width = 24, fig.height= 24, message=FALSE, warning=FALSE,cache=TRUE}
m3 <- movies%>%select(c(imdb_rating,studio))

```
```{r, fig.width = 24, fig.height= 24, message=FALSE, warning=FALSE, cache=TRUE}
#m4 <- movies%>%select(c(imdb_rating,title))
ggplot(data = m3,aes(x = studio, y = imdb_rating))+geom_boxplot()+coord_flip()
##ggplot(data = m4[1:10,], aes(x = title, y = imdb_rating))+geom_boxplot()

```
Let us do an ANOVA to check if there is any relationship, as well as check the adjusted R-squared via a linear model.
```{r}
AnoV <- aov(imdb_rating~studio, data = m3)
summary(AnoV)

lm4 <- lm(imdb_rating~studio, data = m3)
summary(lm4)$r.squared
summary(lm4)$adj.r.squared
```


- boxplot of imdb_rating versus studio shows an interesting variation among the different studios.  Certain studios show a skewed distribution with hign medians, with long left or right skewes, certain other studions have a median high imdb_rating, compared to the rest. This box plot has lots of interesting information. And can be it's own topic of research.

- <span style="color: red"> **NOTE** The difference between R-squared of: 0.4059  and adjusted R-squared: 0.1171. Which is very steep. This seems to indicate that adding studio to our model, might actually reduce the overall adjusted-R squared of our model. And it also seems to show that studio is not necessarily a good predictor variable.</span> 


### <span style="color: Brown">**Result: **</span> 

- We see a high degree of variance between the studios with respect to the IMDB_ratings.  So it looks like studios might have a strong influence in the imdb_ratings.  

- We need to keep this in mind when we build our model.


~~I am interested if the title of a movie, has any influence on audience, affecting the imdb_rating of a movie.~~  

- imdb_url and rt_url, offer no real value, as they merely are locations in the internet of where information about the movies is found.


### <span style="color: Orange">**Step5 : **</span>  

#### <span style="color: darkgreen">**Influence on lead actor on IMDB_ratings:** </span>

- Another factor of interest, is to see if actors have any influence on movie quality thereby influencing the IMDB_rating


- Lead actors can sometimes have a major influence on ratings of movies. I will be exploring the relationship between the **actor1** variable and the **IMDB_ratings** next. 

```{r, fig.width = 24, fig.height= 24, message=FALSE, warning=FALSE, cache=TRUE}
m5 <- movies%>%select(c(imdb_rating,actor1))
ggplot(data = m5,aes(x = actor1, y = imdb_rating))+geom_boxplot()+coord_flip()

```

Let us do a more formal check with an ANOVA analysis as well as fit to a linear model just to make sure. Before we get the R squared value.
```{r}
AnoV <- aov(imdb_rating~actor1, data = m5)
summary(AnoV)

lm5 <- lm(imdb_rating~actor1, data = m5)
summary(lm5)$r.squared
summary(lm5)$adj.r.squared
```


- the boxplot presents lots of interesting information on lead actors and their movies. 

- It says quite a bit about what kind of movies that the actors seem to fall into. for example: **Adam Sandler** although famous seems to be starring in movies with really low IMDB ratings.  **Arnold Schwazneggar** seems to star in movies with overall higher median IMDB rating.

- Although we see variance, can it be completely attributed to just the actor or to other factors such as the director.

- Nevertheless, this is an interestng variable to consider too.

- <span style="color: red"> **NOTE** One really interesting trend is the difference between R-squared of: 0.8342  and adjusted R-squared: 0.3452.  From this we see that adding in more actors into the predictors actually results in a higher penality. So from this it looks like, actor1 is not necessarily a good predictor variable</span>

* * *

## Part 4: Modeling

- We will now begin creating the model by using the forward selection R-squared approach.  

- We will add one predictor variable at a time, and check if the R-squared value increases or decreases.  Predictor variables that lower R-squared will be dropped.

- Our strategy for adding the predictor variables will follow this strategy:-

- First we will be adding the numeric predictor variables, we will add predictors that result in increased adjustd R-squared. We will then run tests on them to make sure that they satisfy the conditions needed to be added in a MLR model. If they fail to satisfy the conditions, we will remove them from the model.

- We will follow the same approach and add in the categorical variables.

#### <span style="color: red">  Let us list out their R-squared and Adjusted R-square values: </span>

```{r, include=FALSE}
predictorVariable <- NULL
rsquared <- NULL
adjrsquared <- NULL
variableType <- NULL

lm2_1 <- lm(imdb_rating~runtime, data = m1)
predictorVariable <-append(predictorVariable,"runtime",after = length(predictorVariable))
variableType <- append(variableType,"Numeric",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))



lm2_1 <- lm(imdb_rating~audience_score, data = m1)
predictorVariable <-append(predictorVariable,"audience_score",after = length(predictorVariable))
variableType <- append(variableType,"Numeric",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))

lm2_1 <- lm(imdb_rating~critics_score, data = m1)
predictorVariable <-append(predictorVariable,"critics_score",after = length(predictorVariable))
variableType <- append(variableType,"Numeric",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))

lm2_1 <- lm(imdb_rating~imdb_num_votes, data = m1)
predictorVariable <-append(predictorVariable,"imdb_num_votes",after = length(predictorVariable))
variableType <- append(variableType,"Numeric",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))


lm2_1 <- lm(imdb_rating~title_type, data = m2)
predictorVariable <-append(predictorVariable,"title_type",after = length(predictorVariable))
variableType <- append(variableType,"Categorical",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))


lm2_1 <- lm(imdb_rating~genre, data = m2)
predictorVariable <-append(predictorVariable,"genre",after = length(predictorVariable))
variableType <- append(variableType,"Categorical",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))


lm2_1 <- lm(imdb_rating~mpaa_rating, data = m2)
predictorVariable <-append(predictorVariable,"mpaa_rating",after = length(predictorVariable))
variableType <- append(variableType,"Categorical",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))


lm2_1 <- lm(imdb_rating~critics_rating, data = m2)
predictorVariable <-append(predictorVariable,"critics_rating",after = length(predictorVariable))
variableType <- append(variableType,"Categorical",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))


lm2_1 <- lm(imdb_rating~audience_rating, data = m2)
predictorVariable <-append(predictorVariable,"audience_rating",after = length(predictorVariable))
variableType <- append(variableType,"Categorical",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))


lm2_1 <- lm(imdb_rating~best_pic_nom, data = m2)
predictorVariable <-append(predictorVariable,"best_pic_nom",after = length(predictorVariable))
variableType <- append(variableType,"Categorical",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))


lm2_1 <- lm(imdb_rating~best_pic_win, data = m2)
predictorVariable <-append(predictorVariable,"best_pic_win",after = length(predictorVariable))
variableType <- append(variableType,"Categorical",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))


lm2_1 <- lm(imdb_rating~best_actor_win, data = m2)
predictorVariable <-append(predictorVariable,"best_actor_win",after = length(predictorVariable))
variableType <- append(variableType,"Categorical",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))


lm2_1 <- lm(imdb_rating~best_actress_win, data = m2)
predictorVariable <-append(predictorVariable,"best_actress_win",after = length(predictorVariable))
variableType <- append(variableType,"Categorical",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))


lm2_1 <- lm(imdb_rating~best_dir_win, data = m2)
predictorVariable <-append(predictorVariable,"best_dir_win",after = length(predictorVariable))
variableType <- append(variableType,"Categorical",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))


lm2_1 <- lm(imdb_rating~top200_box, data = m2)
predictorVariable <-append(predictorVariable,"top200_box",after = length(predictorVariable))
variableType <- append(variableType,"Categorical",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))

lm2_1 <- lm(imdb_rating~studio, data = m3)
predictorVariable <-append(predictorVariable,"studio",after = length(predictorVariable))
variableType <- append(variableType,"Categorical",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))

lm2_1 <- lm(imdb_rating~actor1, data = m5)
predictorVariable <-append(predictorVariable,"actor1",after = length(predictorVariable))
variableType <- append(variableType,"Categorical",after = length(variableType))
rsquared <- append(rsquared,summary(lm2_1)$r.squared,after = length(rsquared))
adjrsquared <- append(adjrsquared,summary(lm2_1)$adj.r.squared,after = length(adjrsquared))

RsquaredSummary <- data.frame( predictorVariable = predictorVariable,VariableType = variableType ,rsquared = rsquared, adjrsquared = adjrsquared)
RsquaredSummary <- RsquaredSummary%>%group_by(VariableType)%>%arrange(desc(adjrsquared), .by_group = TRUE)

```


```{r}
kable(RsquaredSummary)
```

#### <span style="color: darkgreen">**Section 1: Build model with numeric variables** </span>  

- We will be considering the following numeric_variables, since they have the largest correleation coefficients: **audience_score,   critics_score,   imdb_num_votes,   runtime**  

```{r}

model <- lm(imdb_rating~audience_score+critics_score, data = movies)

summary(model)$r.squared
summary(model)$adj.r.squared

# p1 <- ggplot(data = model, aes(x = audience_score, y = .resid))+geom_point()+geom_hline(yintercept=0, linetype="dashed", color = "red")
# 
# 
# p2 <- ggplot(data = model, aes(x = .resid)) + geom_histogram() + xlab("Residuals")
# #QQ-plot
# p2_1 <- ggplot(data = model, aes(sample = .resid)) + stat_qq()+stat_qq_line(size=1.25, color="red")
# 
# p3 <- ggplot(data = model, aes(x = .fitted, y = .resid))+geom_point()+geom_hline(yintercept=0, linetype="dashed", color = "red")
# p3_1 <- ggplot(data = model, aes(x = abs(.fitted), y = .resid))+geom_point()+geom_hline(yintercept=0, linetype="dashed", color = "red")
# 
# 
# print(p1)
# print(p2)
# print(p2_1)
# print(p3)
# print(p3_1)


#hist(model$residuals)
#qqnorm(model$residuals)
#qqline(model$residuals)
#plot(model$residuals~model$fitted)+abline(h = 0)
#plot(abs(model$residuals)~model$fitted)+abline(h = 0)
#plot(model$residuals)+abline(h = 0)
#modelSummary <- data.frame( modelPredVar = modelPredVar,modelRsquare = modelRsquare ,modeladjRsquare = modeladjRsquare)

summary(model)
```

#### <span style="color: red">**Let us plot all the numeric variables to check for costant variability vs residuals** </span> 

```{r}
plot(model$residuals~movies$audience_score)+abline(h = 0)
plot(model$residuals~movies$critics_score)+abline(h = 0)
plot(model$residuals~movies$imdb_num_votes)+abline(h = 0)
plot(model$residuals~movies$runtime)+abline(h = 0)
```

- **audience_score** is constatnly variable , so we keep it.

- **critics_score** is constantly variable so we keep it.

- **imdb_num_votes** is not constantly variable so we drop it.

- **runtime** is not constantly variable so we drop it.

```{r}
model <- lm(imdb_rating~audience_score, data = movies)

```

After adding: **audience_score**  

the R-squared: `r summary(model)$r.squared` 

the adjusted R-squared: `r summary(model)$adj.r.squared`

```{r}
model <- lm(imdb_rating~audience_score+critics_score, data = movies)
```
After adding: **critics_score**  

the R-squared: `r summary(model)$r.squared` 

the adjusted R-squared: `r summary(model)$adj.r.squared`

**The adjusted R-squared has increased, so critics score is a useful predictor variable, along with audience_score**

#### <span style="color: red">**Let us begin adding in the categorical variables now** </span> 

#### <span style="color: red">**audience_rating** </span>
```{r, fig.width = 48, fig.height = 48}
model <- lm(imdb_rating~audience_score+critics_score+audience_rating, data = movies)
par(mfrow=c(2,2))
hist(model$residuals)
qqnorm(model$residuals)
qqline(model$residuals)
plot(model$residuals~model$fitted)+abline(h = 0)
plot(abs(model$residuals)~model$fitted)+abline(h = 0)
plot(model$residuals)+abline(h = 0)
dev.off()
```

- <span style="color: darkgreen">From the plots: audience_rating shows: Constant variability.  Nearly Normal residuals.</span>

After adding: **audience_rating**  

the R-squared: `r summary(model)$r.squared` 

the adjusted R-squared: `r summary(model)$adj.r.squared`

**The adjusted- R squared increased so we keep this predictor**


#### <span style="color: red">**critics_rating** </span>
```{r, fig.width = 48, fig.height = 48}
model <- lm(imdb_rating~audience_score+critics_score+audience_rating+critics_rating, data = movies)
par(mfrow=c(2,2))
hist(model$residuals)
qqnorm(model$residuals)
qqline(model$residuals)
plot(model$residuals~model$fitted)+abline(h = 0)
plot(abs(model$residuals)~model$fitted)+abline(h = 0)
plot(model$residuals)+abline(h = 0)
dev.off()
```

- <span style="color: darkgreen">From the plots: audience_rating shows: Constant variability.  Nearly Normal residuals.</span>

After adding: **critics_rating**  

the R-squared: `r summary(model)$r.squared` 

the adjusted R-squared: `r summary(model)$adj.r.squared`

**The adjusted- R squared increased so we keep this predictor**


#### <span style="color: red">**genre** </span>
```{r, fig.width = 48, fig.height = 48}
model <- lm(imdb_rating~audience_score+critics_score+audience_rating+critics_rating+genre, data = movies)
par(mfrow=c(2,2))
hist(model$residuals)
qqnorm(model$residuals)
qqline(model$residuals)
plot(model$residuals~model$fitted)+abline(h = 0)
plot(abs(model$residuals)~model$fitted)+abline(h = 0)
plot(model$residuals)+abline(h = 0)
dev.off()
```


- <span style="color: darkgreen">From the plots: audience_rating shows: Constant variability.  Nearly Normal residuals.</span>

After adding: **genre**  

the R-squared: `r summary(model)$r.squared` 

the adjusted R-squared: `r summary(model)$adj.r.squared`

**The adjusted- R squared increased so we keep this predictor**



#### <span style="color: red">**best_pic_win** </span>
```{r, fig.width = 48, fig.height = 48}
model <- lm(imdb_rating~audience_score+critics_score+audience_rating+critics_rating+genre+best_pic_win, data = movies)
par(mfrow=c(2,2))
hist(model$residuals)
qqnorm(model$residuals)
qqline(model$residuals)
plot(model$residuals~model$fitted)+abline(h = 0)
plot(abs(model$residuals)~model$fitted)+abline(h = 0)
plot(model$residuals)+abline(h = 0)
dev.off()
```

- <span style="color: darkgreen">From the plots: audience_rating shows: Constant variability.  Nearly Normal residuals.</span>

After adding: **best_pic_win**  

the R-squared: `r summary(model)$r.squared` 

the adjusted R-squared: `r summary(model)$adj.r.squared`

**The adjusted- R squared increased so we keep this predictor**



- The remaining categorical variables were ignored as they caused a decrease in adjuster - r squared.

- Also studio and actor1 are not really categorical variables, as they do not have categories , but just a whole bunch of actor and studio names.

- There fore in our model they are not used. 

* * *

## Part 5: Prediction

#### <span style="color: red">  Wonder Woman(2017)  </span>
```{r}
#Wonder Woman rating
newprof <- data.frame(audience_score = 87 ,critics_score = 93 ,audience_rating = "Upright" ,critics_rating = "Certified Fresh",genre = "Action & Adventure",best_pic_win = "no")
p1 <- predict(model, newprof, interval = "prediction", level = 0.95)
kable(p1)
```

<span style="color: red"> predicted value: </span> `r p1[1]` 

<span style="color: red"> original IMDB rating: </span> [7.4](https://www.imdb.com/title/tt0451279/) 

<span style="color: red"> prediction off by(percentage): </span>  `r ((p1[1]-7.4)/10)*100` 

<span style="color: darkgreen"> ORIGINAL IMDB VALUE falls inside PREDICTED Confidence interval of 95% </span> 


#### <span style="color: red">  Star Wars: The Last Jedi (2018)  </span>

```{r}
#SW The Last Jedi
newprof <- data.frame(audience_score = 43 ,critics_score = 91 ,audience_rating = "Spilled" ,critics_rating = "Certified Fresh",genre = "Action & Adventure",best_pic_win = "no")
predict(model, newprof, interval = "prediction", level = 0.95)
p1 <- predict(model, newprof, interval = "prediction", level = 0.95)
kable(p1)
```

<span style="color: red"> predicted value: </span> `r p1[1]` 

<span style="color: red"> original IMDB rating: </span> [7.0](https://www.imdb.com/title/tt2527336/) 

<span style="color: red"> prediction off by(percentage): </span>  `r ((p1[1]-7)/10)*100` 

<span style="color: darkgreen"> ORIGINAL IMDB VALUE falls inside PREDICTED Confidence interval of 95% </span> 



#### <span style="color: red">  Star Wars: The Rise of Skywalker (2019)  </span>

```{r}
#SW The rise of skywalker
newprof <- data.frame(audience_score = 86 ,critics_score = 52 ,audience_rating = "Upright" ,critics_rating = "Rotten",genre = "Action & Adventure",best_pic_win = "no")
predict(model, newprof, interval = "prediction", level = 0.95)
p1 <- predict(model, newprof, interval = "prediction", level = 0.95)
kable(p1)
```

<span style="color: red"> predicted value: </span> `r p1[1]` 

<span style="color: red"> original IMDB rating: </span> [6.7](https://www.imdb.com/title/tt2527338/) 

<span style="color: red"> prediction off by(percentage): </span>  `r ((p1[1]-6.7)/10)*100` 

<span style="color: darkgreen"> ORIGINAL IMDB VALUE falls inside PREDICTED Confidence interval of 95% </span> 


#### <span style="color: red">  James Bond 007: Casino Royale 007  </span>

```{r}
#James Bond 007: Casino Royale 007
newprof <- data.frame(audience_score = 89 ,critics_score = 95 ,audience_rating = "Upright" ,critics_rating = "Certified Fresh",genre = "Action & Adventure",best_pic_win = "no")
predict(model, newprof, interval = "prediction", level = 0.95)
p1 <- predict(model, newprof, interval = "prediction", level = 0.95)
kable(p1)
```


<span style="color: red"> predicted value: </span> `r p1[1]` 

<span style="color: red"> original IMDB rating: </span> [8](https://www.imdb.com/title/tt0381061/) 

<span style="color: red"> prediction off by(percentage): </span>  `r ((p1[1]-8)/10)*100` 

<span style="color: darkgreen"> ORIGINAL IMDB VALUE falls inside PREDICTED Confidence interval of 95% </span>   

* * *

## Part 6: Conclusion

- Our model is **really accurate in the 95% confidence interval** . It is able to predicted IMDB rating values accurately such that they fall into this confidence interval.

- The mean loss of precision of our model is: **`r (abs(((p1[1]-8)/10)*100)+abs(((p1[1]-6.7)/10)*100)+abs(((p1[1]-7)/10)*100)+abs(((p1[1]-7.4)/10)*100))/4`**

- Our model is **well designed** as it takes into account, diveregence of opinion between Critics and Audiences.`

- This is a robust well designed model.
