---
title: "Arizona/US/India Corona Trends- 16th April 2020 with tomorrow's prediction"
author: "Whitchurch Muthumani"
date: "4/16/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Arizona Corona Dashboard

```{r include=FALSE}
library(ggplot2)
library(gridExtra)
library(dplyr)

coronaGrowth <- function(absoluteNumbers){
  growthfactor <- NULL
  absgrowth <- NULL
  for(i in 1:length(absoluteNumbers))
  {
    #print((absoluteNumbers[i]-absoluteNumbers[i+1]/absoluteNumbers[i+1])/absoluteNumbers[i+1])
    growthfactor[i] <- ((absoluteNumbers[i]-absoluteNumbers[i+1]))/(absoluteNumbers[i+1]-absoluteNumbers[i+2])  
    #growthfactor[i] <- ((absoluteNumbers[i]-absoluteNumbers[i+1]))/(absolutenumbers[i])
  }
  
  return (rev(growthfactor))
}

DailyGrowthRate <- function(absoluteNumbers)
{
  growthfactor <- NULL
  absgrowth <- NULL
  for(i in 1:length(absoluteNumbers))
  {
    #print((absoluteNumbers[i]-absoluteNumbers[i+1]/absoluteNumbers[i+1])/absoluteNumbers[i+1])
    growthfactor[i] <- ((absoluteNumbers[i]/absoluteNumbers[i+1]))-1 
    #growthfactor[i] <- ((absoluteNumbers[i]-absoluteNumbers[i+1]))/(absolutenumbers[i])
  }
  
  return (rev(growthfactor))  
}

DoublingRate <- function(absoluteNumbers)
{
  doublingrRate <- NULL
  cumsum <- NULL
  for(i in 1:length(absoluteNumbers))
  {
    
    doublingrRate[i] <- (70/(absoluteNumbers[i]*100))
    
  }
  return ((doublingrRate))  
}

#Calculate the mean and median growth rate:
MeanMediangrowthrate <- function(dataTrend1)
{
  MeanGRTrend <- NULL
  MedianGRTrend <- NULL
  
  for(i in 1:length(dataTrend1$GrowthRate))
  {
    MedianGRTrend[i] <- median(dataTrend1[1:i,"GrowthRate"])
    MeanGRTrend[i] <- mean(dataTrend1[1:i,"GrowthRate"])
  }
  
  dataTrend1$MedianGRT = MedianGRTrend
  dataTrend1$MeanGRT = MeanGRTrend
  
  return(dataTrend1)

}

#Calculate ratio of newdeaths/newcases
newDeathstonewCases <- function(absolutenumbers,totaldeaths)
{
  newCases <- NULL
  newDeaths <- NULL
  for(i in 1:length(absolutenumbers))
  {
    if(i == length(absolutenumbers))
    {
      print("reached the end")
    }else{
     newDeaths[i] = (totaldeaths[i]-totaldeaths[i+1])
     newCases[i] = (absolutenumbers[i]-absolutenumbers[i+1]) 
    }
  }
  

  days <- seq(2:length(absolutenumbers))
 return(data.frame( newDeaths = rev(newDeaths),newCases = rev(newCases),Days = days))
 
}

#Calculate ratio of deaths to totalCases
totalDeathstoCases <- function(absolutenumbers,totaldeaths)
{
  ratio <- NULL
  for(i in 1:length(absolutenumbers))
  {

     ratio[i] = (totaldeaths[i])/(absolutenumbers[i]) 
  }
  
  return(rev(ratio))
}
```


```{r include=FALSE}
#Arizona

absolutenumbers <- c(4234,3962,3806,3702,3539,3393,3112,3018,2726,2575,2456,2269,2019,1769,1598,1413,1289,1157,919,773,665,508,401,326,234,152,104,71,44,28,20,18,12)

totaldeaths <- c(150,142,131,122,115,108,97,89,80,73,65,64,52,41,32,29,24,20,17,15,13,9,6,6,3,2,2,1,0,0,0,0,0)

#America
#absolutenumbers <- c(101.7*1000,83.8*1000,65.8*1000,53.7*1000,43.8*1000,33.3*1000,25*1000,19*1000,13.7*1000,7.8*1000,6.4*1000,4.6*1000,3.5*1000,2.7*1000)
```

```{r include=FALSE,message=FALSE}
growthrate <- na.exclude(coronaGrowth(absolutenumbers))
days <- seq(1:length(growthrate))
dataTrend <- data.frame(GrowthRate = growthrate,Days = days)
print(dataTrend)
p1 <- ggplot(data=dataTrend, aes(x=Days, y=GrowthRate, group=1)) +ggtitle("Arizona: Rate of change in Growth Rate")+
  geom_line()+
  geom_point()

print(dataTrend)
#print(dataTrend1)




growthrate1 <- na.exclude(DailyGrowthRate(absolutenumbers))
doublingrate1 <- na.exclude(DoublingRate(growthrate1))
days1 <- seq(1:length(growthrate1))
dataTrend1 <- data.frame(GrowthRate = growthrate1,Days = days1,doublinrRate = doublingrate1)
print(dataTrend1)
multiplier <- tail(dataTrend1, n = 1)
pred_val <- (head(absolutenumbers, n = 1)*multiplier[,"GrowthRate"])+head(absolutenumbers, n = 1)

p2 <- ggplot(data=dataTrend1, aes(x=Days, y=GrowthRate, group=1)) +ggtitle("Arizona: Growth Rate")+
  geom_line()+
  geom_point()

days <- seq(1:length(absolutenumbers))
growthrate <- append(growthrate, c(1,1),after = 0)
dataOriginal <- data.frame(Cases = rev(absolutenumbers), Days = days)
p3<- ggplot() +geom_line(data=dataOriginal, aes(x=Days, y=Cases,),color = "Red")+labs(title="Arizona: Daily Cases", y="Cases", x="Days")

p3_log<- ggplot() +geom_line(data=dataOriginal, aes(x=Days, y=log2(Cases)),color = "Red")+labs(title="Arizona: Daily Cases", y="Cases", x="Days")+
  #geom_line(data=dataOriginal, aes(x=Days+1, y=Prediction),color = "Blue")
  geom_point()



# Analyze growth rate to get interval of values
summary(dataTrend1$GrowthRate)
dataTrend1%>%select("GrowthRate")%>%summarise(meanGR = mean(GrowthRate), medianGR = median(GrowthRate), sdGR = sd(GrowthRate))
p4 <- ggplot(data = dataTrend1, aes(x = GrowthRate))+ggtitle("Histogram of GrowthRate")+geom_histogram(binwidth = 0.01)


#Analyze the doubling rate
p5<- ggplot() +geom_line(data=dataTrend1, aes(x=Days, y=doublinrRate),color = "Red")+labs(title="Doubling Rate in days", y="Days to double", x ="per day estimate")+
  #geom_line(data=dataOriginal, aes(x=Days+1, y=Prediction),color = "Blue")
  geom_point()

#Scatter plor of Growthrate versus Doubling Rate
p6 <- ggplot()+geom_point(data = dataTrend1, aes(x = doublinrRate, y = GrowthRate))+labs(title = "Doubling Rate vs Growth Rate")

dataTrend1 <- MeanMediangrowthrate(dataTrend1 = dataTrend1)

#Median growth rate trend 
#Median growth rate trend 
p7 <- ggplot()+geom_line(data = dataTrend1, aes(x = Days, y = MedianGRT),color = "Blue")+labs(title = "Arizona Median(BLUE)VsMean(RED) Growth Rate")+geom_line(data=dataTrend1, aes(x=Days , y=MeanGRT),color = "Red")


deathrate1 <- na.exclude(DailyGrowthRate(totaldeaths))
doublingdeathrate1 <- na.exclude(DoublingRate(deathrate1))
days1 <- seq(1:length(deathrate1))
dataDeathTrend1 <- data.frame(DeathRate = deathrate1,Days = days1,doublinrRate = doublingdeathrate1)
#print(dataDeathTrend1)
#multiplier <- tail(totaldeaths, n = 1)
#pred_val <- (head(dataDeathTrend1, n = 1)*multiplier[,"DeathRate"])+head(dataDeathTrend1, n = 1)

p8 <- ggplot(data=dataDeathTrend1, aes(x=Days, y=log2(DeathRate), group=1)) +ggtitle("Arizona: Daily Death Rate")+
  geom_line(color = "Green")+
  geom_point()


rationewdeathtocases <- newDeathstonewCases(absolutenumbers,totaldeaths)

p9 <- ggplot(data=rationewdeathtocases, aes(x=Days, y=log2(newCases), group=1)) +ggtitle("Arizona:newCases(Orange)")+
  geom_line(color = "Orange")+geom_point()

totaldeathstocases <- totalDeathstoCases(absolutenumbers,totaldeaths)
days <- seq(1:length(absolutenumbers))
datatotaldeathstocasesTrend <- data.frame(ratio = totaldeathstocases,Days = days)
#Scatter plot of totalcases to totaldeaths
p10 <- ggplot(data=datatotaldeathstocasesTrend, aes(x=Days, y=ratio, group=1)) +ggtitle("Arizona: TotalDeaths/TotalCases")+
  geom_line(color = "Yellow")+
  geom_point()

#Confidence Interval for mean population GrowthRate:
meanGR = mean(dataTrend1$GrowthRate)
sdGR = sd(dataTrend1$GrowthRate)

uppperlimitGR= meanGR+sdGR
lowerlimitGR = meanGR-sdGR

#Predictions based on: Current Growth Rate, MeanGrowth Rate, Upper GrowthRate, Lower GrowthRate
pred_val_Current <- (head(absolutenumbers, n = 1)*multiplier[,"GrowthRate"])+head(absolutenumbers, n = 1)
pred_val_MeanGR <- (head(absolutenumbers, n = 1)*meanGR)+head(absolutenumbers, n = 1)
pred_val_upperGR <- (head(absolutenumbers, n = 1)*uppperlimitGR)+head(absolutenumbers, n = 1)
pred_val_lowerGR <- (head(absolutenumbers, n = 1)*lowerlimitGR)+head(absolutenumbers, n = 1)

print("Current GR:")
print(multiplier[,"GrowthRate"])
print(pred_val_Current)

print("Mean GR:")
print(meanGR)
print(pred_val_MeanGR)

print("Upperlimit GR:")
print(uppperlimitGR)
print(pred_val_upperGR)

print("Lowerlimit GR:")
print(lowerlimitGR)
print(pred_val_lowerGR)

#Doubling rate:-
DRCurrent <- round(70/(meanGR*100),digits = 0)
DRMedian <- round(70/(median(dataTrend1$GrowthRate)*100),digits = 0)
latestDRCurrent <- round(tail(dataTrend1$doublinrRate,n=1),digits = 0)

```

```{r echo=FALSE,warning=FALSE}
#grid.arrange(p3,p3_log,p1,p2,p4,p5,p6,nrow = 4, ncol = 2)
print(p3)
print(p3_log)
print(p1)
print(p2)
print(p4)
print(p5)
print(p6)
print(p8)
print(p9)
print(p10)
print(p7)
```

### Results:
- Tomorrows predicted case number is between **`r pred_val_Current`** and **`r pred_val_MeanGR`** cases.
- Current doubling rate is **`r latestDRCurrent` days**
- Mean doubling rate, i.e number of cases doubles every **`r DRCurrent` days**.
- Median doubling rate, i.e number of cases doubles every **`r DRMedian` days**.
- Latest Observed Growth Rate **`r multiplier[,"GrowthRate"]*100`** percent.
- Mean Growth Rate: **`r meanGR*100`** percent.
- Median Growth Rate: **`r median(dataTrend1$GrowthRate)*100`** percent.
- Standard deviation: +/- **`r sdGR*100`**.



### US Corona Dashboard

```{r include=FALSE}
library(ggplot2)
library(gridExtra)
library(dplyr)

coronaGrowth <- function(absoluteNumbers){
  growthfactor <- NULL
  absgrowth <- NULL
  for(i in 1:length(absoluteNumbers))
  {
    #print((absoluteNumbers[i]-absoluteNumbers[i+1]/absoluteNumbers[i+1])/absoluteNumbers[i+1])
    growthfactor[i] <- ((absoluteNumbers[i]-absoluteNumbers[i+1]))/(absoluteNumbers[i+1]-absoluteNumbers[i+2])  
    #growthfactor[i] <- ((absoluteNumbers[i]-absoluteNumbers[i+1]))/(absolutenumbers[i])
  }
  
  return (rev(growthfactor))
}

DailyGrowthRate <- function(absoluteNumbers)
{
  growthfactor <- NULL
  absgrowth <- NULL
  for(i in 1:length(absoluteNumbers))
  {
    #print((absoluteNumbers[i]-absoluteNumbers[i+1]/absoluteNumbers[i+1])/absoluteNumbers[i+1])
    growthfactor[i] <- ((absoluteNumbers[i]/absoluteNumbers[i+1]))-1 
    #growthfactor[i] <- ((absoluteNumbers[i]-absoluteNumbers[i+1]))/(absolutenumbers[i])
  }
  
  return (rev(growthfactor))  
}

DoublingRate <- function(absoluteNumbers)
{
  doublingrRate <- NULL
  cumsum <- NULL
  for(i in 1:length(absoluteNumbers))
  {
    
    doublingrRate[i] <- (70/(absoluteNumbers[i]*100))
    
  }
  return ((doublingrRate))  
}

#Calculate the mean and median growth rate:
MeanMediangrowthrate <- function(dataTrend1)
{
  MeanGRTrend <- NULL
  MedianGRTrend <- NULL
  
  for(i in 1:length(dataTrend1$GrowthRate))
  {
    MedianGRTrend[i] <- median(dataTrend1[1:i,"GrowthRate"])
    MeanGRTrend[i] <- mean(dataTrend1[1:i,"GrowthRate"])
  }
  
  dataTrend1$MedianGRT = MedianGRTrend
  dataTrend1$MeanGRT = MeanGRTrend
  
  return(dataTrend1)

}

# #Calculate ratio of newdeaths/newcases
# newDeathstonewCases <- function(absolutenumbers,totaldeaths)
# {
#   ratio <- NULL
#   for(i in 1:length(absolutenumbers))
#   {
#     if(i == length(absolutenumbers))
#     {
#       print("reached the end")
#     }else{
#      ratio[i] = (totaldeaths[i]-totaldeaths[i+1])/(absolutenumbers[i]-absolutenumbers[i+1]) 
#     }
#   }
#   
#   return(rev(ratio))
# }

#Calculate ratio of newdeaths/newcases
newDeathstonewCases <- function(absolutenumbers,totaldeaths)
{
  newCases <- NULL
  newDeaths <- NULL
  for(i in 1:length(absolutenumbers))
  {
    if(i == length(absolutenumbers))
    {
      print("reached the end")
    }else{
     newDeaths[i] = (totaldeaths[i]-totaldeaths[i+1])
     newCases[i] = (absolutenumbers[i]-absolutenumbers[i+1]) 
    }
  }
  

  days <- seq(2:length(absolutenumbers))
 return(data.frame( newDeaths = rev(newDeaths),newCases = rev(newCases),Days = days))
 
}

#Calculate ratio of deaths to totalCases
totalDeathstoCases <- function(absolutenumbers,totaldeaths)
{
  ratio <- NULL
  for(i in 1:length(absolutenumbers))
  {

     ratio[i] = (totaldeaths[i])/(absolutenumbers[i]) 
  }
  
  return(rev(ratio))
}

```


```{r include=FALSE}
#Arizona
#absolutenumbers <- c(773,665,508,401,326,234,152,104,71,44,28,20,18,12)

#America

absolutenumbers <- c(654.343*1000,622.412*1000,591.285*1000,573.138*1000,545.934*1000,521.623*1000,489.646*1000,454.615*1000,418.451*1000,386.571*1000,355.834*1000,327.871*1000,300.432*1000,265.506*1000,235.281*1000,205.035*1000,176.518*1000,145.369*1000,133.094*1000,123.578*1000,101.7*1000,83.8*1000,65.8*1000,53.7*1000,43.8*1000,33.3*1000,25*1000,19*1000,13.7*1000,7.8*1000,6.4*1000,4.6*1000,3.5*1000,2.7*1000)

totaldeaths <- c(33490,27549,24618,22930,21474,20057,18325,16691,14791,12851,10880,9622,8457,7127,6081,5107,4057,3143,2585,2222,1697,1296,1028,781,556,414,301,255,206,150,110,87,69,57,48,41,38,30,26,22,19,15,12,11,9,6,1,1)




```

```{r include=FALSE,message=FALSE}
growthrate <- na.exclude(coronaGrowth(absolutenumbers))
days <- seq(1:length(growthrate))
dataTrend <- data.frame(GrowthRate = growthrate,Days = days)
print(dataTrend)
p1 <- ggplot(data=dataTrend, aes(x=Days, y=GrowthRate, group=1)) +ggtitle("US: Rate of change in Growth Rate")+
  geom_line()+
  geom_point()

print(dataTrend)
#print(dataTrend1)




growthrate1 <- na.exclude(DailyGrowthRate(absolutenumbers))
doublingrate1 <- na.exclude(DoublingRate(growthrate1))
days1 <- seq(1:length(growthrate1))
dataTrend1 <- data.frame(GrowthRate = growthrate1,Days = days1,doublinrRate = doublingrate1)
print(dataTrend1)
multiplier <- tail(dataTrend1, n = 1)
pred_val <- (head(absolutenumbers, n = 1)*multiplier[,"GrowthRate"])+head(absolutenumbers, n = 1)

p2 <- ggplot(data=dataTrend1, aes(x=Days, y=GrowthRate, group=1)) +ggtitle("US: Growth Rate")+
  geom_line()+
  geom_point()

days <- seq(1:length(absolutenumbers))
growthrate <- append(growthrate, c(1,1),after = 0)
dataOriginal <- data.frame(Cases = rev(absolutenumbers), Days = days)
p3<- ggplot() +geom_line(data=dataOriginal, aes(x=Days, y=Cases,),color = "Red")+labs(title="US: Daily Cases", y="Cases", x="Days")

p3_log<- ggplot() +geom_line(data=dataOriginal, aes(x=Days, y=log2(Cases)),color = "Red")+labs(title="US: Daily Cases", y="Cases", x="Days")+
  #geom_line(data=dataOriginal, aes(x=Days+1, y=Prediction),color = "Blue")
  geom_point()



# Analyze growth rate to get interval of values
summary(dataTrend1$GrowthRate)
dataTrend1%>%select("GrowthRate")%>%summarise(meanGR = mean(GrowthRate), medianGR = median(GrowthRate), sdGR = sd(GrowthRate))
p4 <- ggplot(data = dataTrend1, aes(x = GrowthRate))+ggtitle("Histogram of GrowthRate")+geom_histogram(binwidth = 0.01)


#Analyze the doubling rate
p5<- ggplot() +geom_line(data=dataTrend1, aes(x=Days, y=doublinrRate),color = "Red")+labs(title="Doubling Rate in days", y="Days to double", x ="per day estimate")+
  #geom_line(data=dataOriginal, aes(x=Days+1, y=Prediction),color = "Blue")
  geom_point()

#Scatter plor of Growthrate versus Doubling Rate
p6 <- ggplot()+geom_point(data = dataTrend1, aes(x = doublinrRate, y = GrowthRate))+labs(title = "Doubling Rate vs Growth Rate")

dataTrend1 <- MeanMediangrowthrate(dataTrend1 = dataTrend1)


#Median growth rate trend 
#Median growth rate trend 
p7 <- ggplot()+geom_line(data = dataTrend1, aes(x = Days, y = MedianGRT),color = "Blue")+labs(title = "US Median(BLUE)VsMean(RED) Growth Rate")+geom_line(data=dataTrend1, aes(x=Days , y=MeanGRT),color = "Red")

deathrate1 <- na.exclude(DailyGrowthRate(totaldeaths))
doublingdeathrate1 <- na.exclude(DoublingRate(deathrate1))
days1 <- seq(1:length(deathrate1))
dataDeathTrend1 <- data.frame(DeathRate = deathrate1,Days = days1,doublinrRate = doublingdeathrate1)
#print(dataDeathTrend1)
#multiplier <- tail(totaldeaths, n = 1)
#pred_val <- (head(dataDeathTrend1, n = 1)*multiplier[,"DeathRate"])+head(dataDeathTrend1, n = 1)

p8 <- ggplot(data=dataDeathTrend1, aes(x=Days, y=log2(DeathRate), group=1)) +ggtitle("US: Daily Death Rate")+
  geom_line(color = "Green")+
  geom_point()

rationewdeathtocases <- newDeathstonewCases(absolutenumbers,totaldeaths)

p9 <- ggplot(data=rationewdeathtocases, aes(x=Days, y=log2(newCases), group=1)) +ggtitle("US:newCases(Orange)")+
  geom_line(color = "Orange")+geom_point()
totaldeathstocases <- totalDeathstoCases(absolutenumbers,totaldeaths)
days <- seq(1:length(absolutenumbers))
datatotaldeathstocasesTrend <- data.frame(ratio = totaldeathstocases,Days = days)
#Scatter plot of totalcases to totaldeaths
p10 <- ggplot(data=datatotaldeathstocasesTrend, aes(x=Days, y=ratio, group=1)) +ggtitle("US: TotalDeaths/TotalCases")+
  geom_line(color = "Yellow")+
  geom_point()
#Confidence Interval for mean population GrowthRate:
meanGR = mean(dataTrend1$GrowthRate)
sdGR = sd(dataTrend1$GrowthRate)

uppperlimitGR= meanGR+sdGR
lowerlimitGR = meanGR-sdGR

#Predictions based on: Current Growth Rate, MeanGrowth Rate, Upper GrowthRate, Lower GrowthRate
pred_val_Current <- (head(absolutenumbers, n = 1)*multiplier[,"GrowthRate"])+head(absolutenumbers, n = 1)
pred_val_MeanGR <- (head(absolutenumbers, n = 1)*meanGR)+head(absolutenumbers, n = 1)
pred_val_upperGR <- (head(absolutenumbers, n = 1)*uppperlimitGR)+head(absolutenumbers, n = 1)
pred_val_lowerGR <- (head(absolutenumbers, n = 1)*lowerlimitGR)+head(absolutenumbers, n = 1)

print("Current GR:")
print(multiplier[,"GrowthRate"])
print(pred_val_Current)

print("Mean GR:")
print(meanGR)
print(pred_val_MeanGR)

print("Upperlimit GR:")
print(uppperlimitGR)
print(pred_val_upperGR)

print("Lowerlimit GR:")
print(lowerlimitGR)
print(pred_val_lowerGR)

#Doubling rate:-
DRCurrent <- round(70/(meanGR*100),digits = 0)
DRMedian <- round(70/(median(dataTrend1$GrowthRate)*100),digits = 0)
latestDRCurrent <- round(tail(dataTrend1$doublinrRate,n=1),digits = 0)

```

```{r echo=FALSE,warning=FALSE}
#grid.arrange(p3,p3_log,p1,p2,p4,p5,p6,nrow = 4, ncol = 2)
print(p3)
print(p3_log)
print(p1)
print(p2)
print(p4)
print(p5)
print(p6)
print(p8)
print(p9)
print(p10)
print(p7)
```

### Results:
- Tomorrows predicted case number is between **`r pred_val_Current`** and **`r pred_val_MeanGR`** cases.
- Current doubling rate is **`r latestDRCurrent` days**
- Mean doubling rate, i.e number of cases doubles every **`r DRCurrent` days**.
- Median doubling rate, i.e number of cases doubles every **`r DRMedian` days**.
- Latest Observed Growth Rate **`r multiplier[,"GrowthRate"]*100`** percent.
- Mean Growth Rate: **`r meanGR*100`** percent.
- Median Growth Rate: **`r median(dataTrend1$GrowthRate)*100`** percent.
- Standard deviation: +/- **`r sdGR*100`**.

### India Corona Dashboard

```{r include=FALSE}
library(ggplot2)
library(gridExtra)
library(dplyr)

coronaGrowth <- function(absoluteNumbers){
  growthfactor <- NULL
  absgrowth <- NULL
  for(i in 1:length(absoluteNumbers))
  {
    #print((absoluteNumbers[i]-absoluteNumbers[i+1]/absoluteNumbers[i+1])/absoluteNumbers[i+1])
    growthfactor[i] <- ((absoluteNumbers[i]-absoluteNumbers[i+1]))/(absoluteNumbers[i+1]-absoluteNumbers[i+2])  
    #growthfactor[i] <- ((absoluteNumbers[i]-absoluteNumbers[i+1]))/(absolutenumbers[i])
  }
  
  return (rev(growthfactor))
}

DailyGrowthRate <- function(absoluteNumbers)
{
  growthfactor <- NULL
  absgrowth <- NULL
  for(i in 1:length(absoluteNumbers))
  {
    #print((absoluteNumbers[i]-absoluteNumbers[i+1]/absoluteNumbers[i+1])/absoluteNumbers[i+1])
    growthfactor[i] <- ((absoluteNumbers[i]/absoluteNumbers[i+1]))-1 
    #growthfactor[i] <- ((absoluteNumbers[i]-absoluteNumbers[i+1]))/(absolutenumbers[i])
  }
  
  return (rev(growthfactor))  
}

DoublingRate <- function(absoluteNumbers)
{
  doublingrRate <- NULL
  cumsum <- NULL
  for(i in 1:length(absoluteNumbers))
  {
    
    doublingrRate[i] <- (70/(absoluteNumbers[i]*100))
    
  }
  return ((doublingrRate))  
}

#Calculate the mean and median growth rate:
MeanMediangrowthrate <- function(dataTrend1)
{
  MeanGRTrend <- NULL
  MedianGRTrend <- NULL
  
  for(i in 1:length(dataTrend1$GrowthRate))
  {
    MedianGRTrend[i] <- median(dataTrend1[1:i,"GrowthRate"])
    MeanGRTrend[i] <- mean(dataTrend1[1:i,"GrowthRate"])
  }
  
  dataTrend1$MedianGRT = MedianGRTrend
  dataTrend1$MeanGRT = MeanGRTrend
  
  return(dataTrend1)

}

#Calculate ratio of newdeaths/newcases
# newDeathstonewCases <- function(absolutenumbers,totaldeaths)
# {
#   ratio <- NULL
#   for(i in 1:length(absolutenumbers))
#   {
#     if(i == length(absolutenumbers))
#     {
#       print("reached the end")
#     }else{
#      ratio[i] = (totaldeaths[i]-totaldeaths[i+1])/(absolutenumbers[i]-absolutenumbers[i+1]) 
#     }
#   }
#   
#   return(rev(ratio))
# }

#Calculate ratio of newdeaths/newcases
newDeathstonewCases <- function(absolutenumbers,totaldeaths)
{
  newCases <- NULL
  newDeaths <- NULL
  for(i in 1:length(absolutenumbers))
  {
    if(i == length(absolutenumbers))
    {
      print("reached the end")
    }else{
     newDeaths[i] = (totaldeaths[i]-totaldeaths[i+1])
     newCases[i] = (absolutenumbers[i]-absolutenumbers[i+1]) 
    }
  }
  

  days <- seq(2:length(absolutenumbers))
 return(data.frame( newDeaths = rev(newDeaths),newCases = rev(newCases),Days = days))
 
}

#Calculate ratio of deaths to totalCases
totalDeathstoCases <- function(absolutenumbers,totaldeaths)
{
  ratio <- NULL
  for(i in 1:length(absolutenumbers))
  {

     ratio[i] = (totaldeaths[i])/(absolutenumbers[i]) 
  }
  
  return(rev(ratio))
}
```


```{r include=FALSE}
#Arizona
#absolutenumbers <- c(773,665,508,401,326,234,152,104,71,44,28,20,18,12)

#America
absolutenumbers <- c(12759,12320,10941,9635,9166,8339,7598,6653,5749,5311,4778,3588,3082,2567,2341,1637,1397,1251,1071,1024,987,887,727,657,536,499,396,332,249,194,169,143,129,114,100,82,74,62,62,47,40,34,31,30,29,7,6,3,3,3,3,3,3,3,3,3,3,3,3,3,3,3)

totaldeaths <- c(423,405,368,331,325,288,246,227,178,160,136,118,99,86,72,58,35,27,24,20,20,12,10,10,7,5,5,4,3,3,2,2,2,2,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
```

```{r include=FALSE,message=FALSE}
growthrate <- na.exclude(coronaGrowth(absolutenumbers))
days <- seq(1:length(growthrate))
dataTrend <- data.frame(GrowthRate = growthrate,Days = days)
print(dataTrend)
p1 <- ggplot(data=dataTrend, aes(x=Days, y=GrowthRate, group=1)) +ggtitle("US: Rate of change in Growth Rate")+
  geom_line()+
  geom_point()

print(dataTrend)
#print(dataTrend1)




growthrate1 <- na.exclude(DailyGrowthRate(absolutenumbers))
doublingrate1 <- na.exclude(DoublingRate(growthrate1))
days1 <- seq(1:length(growthrate1))
dataTrend1 <- data.frame(GrowthRate = growthrate1,Days = days1,doublinrRate = doublingrate1)
print(dataTrend1)
multiplier <- tail(dataTrend1, n = 1)
pred_val <- (head(absolutenumbers, n = 1)*multiplier[,"GrowthRate"])+head(absolutenumbers, n = 1)

p2 <- ggplot(data=dataTrend1, aes(x=Days, y=GrowthRate, group=1)) +ggtitle("India: Growth Rate")+
  geom_line()+
  geom_point()

days <- seq(1:length(absolutenumbers))
growthrate <- append(growthrate, c(1,1),after = 0)
dataOriginal <- data.frame(Cases = rev(absolutenumbers), Days = days)
p3<- ggplot() +geom_line(data=dataOriginal, aes(x=Days, y=Cases,),color = "Red")+labs(title="India: Daily Cases", y="Cases", x="Days")

p3_log<- ggplot() +geom_line(data=dataOriginal, aes(x=Days, y=log2(Cases)),color = "Red")+labs(title="India: Daily Cases", y="Cases", x="Days")+
  #geom_line(data=dataOriginal, aes(x=Days+1, y=Prediction),color = "Blue")
  geom_point()



# Analyze growth rate to get interval of values
summary(dataTrend1$GrowthRate)
dataTrend1%>%select("GrowthRate")%>%summarise(meanGR = mean(GrowthRate), medianGR = median(GrowthRate), sdGR = sd(GrowthRate))
p4 <- ggplot(data = dataTrend1, aes(x = GrowthRate))+ggtitle("Histogram of GrowthRate")+geom_histogram(binwidth = 0.01)


#Analyze the doubling rate
p5<- ggplot() +geom_line(data=dataTrend1, aes(x=Days, y=doublinrRate),color = "Red")+labs(title="Doubling Rate in days", y="Days to double", x ="per day estimate")+
  #geom_line(data=dataOriginal, aes(x=Days+1, y=Prediction),color = "Blue")
  geom_point()

#Scatter plor of Growthrate versus Doubling Rate
p6 <- ggplot()+geom_point(data = dataTrend1, aes(x = doublinrRate, y = GrowthRate))+labs(title = "Doubling Rate vs Growth Rate")

dataTrend1 <- MeanMediangrowthrate(dataTrend1 = dataTrend1)


#Median growth rate trend 
p7 <- ggplot()+geom_line(data = dataTrend1, aes(x = Days, y = MedianGRT),color = "Blue")+labs(title = "India Median(BLUE)VsMean(RED) Growth Rate")+geom_line(data=dataTrend1, aes(x=Days , y=MeanGRT),color = "Red")

deathrate1 <- na.exclude(DailyGrowthRate(totaldeaths))
doublingdeathrate1 <- na.exclude(DoublingRate(deathrate1))
days1 <- seq(1:length(deathrate1))
dataDeathTrend1 <- data.frame(DeathRate = deathrate1,Days = days1,doublinrRate = doublingdeathrate1)
#print(dataDeathTrend1)
#multiplier <- tail(totaldeaths, n = 1)
#pred_val <- (head(dataDeathTrend1, n = 1)*multiplier[,"DeathRate"])+head(dataDeathTrend1, n = 1)

p8 <- ggplot(data=dataDeathTrend1, aes(x=Days, y=log2(DeathRate), group=1)) +ggtitle("India: Daily Death Rate")+
  geom_line(color = "Green")+
  geom_point()

rationewdeathtocases <- newDeathstonewCases(absolutenumbers,totaldeaths)

p9 <- ggplot(data=rationewdeathtocases, aes(x=Days, y=log2(newCases), group=1)) +ggtitle("India:newCases(Orange)")+
  geom_line(color = "Orange")+geom_point()

totaldeathstocases <- totalDeathstoCases(absolutenumbers,totaldeaths)
days <- seq(1:length(absolutenumbers))
datatotaldeathstocasesTrend <- data.frame(ratio = totaldeathstocases,Days = days)
#Scatter plot of totalcases to totaldeaths
p10 <- ggplot(data=datatotaldeathstocasesTrend, aes(x=Days, y=ratio, group=1)) +ggtitle("India: TotalDeaths/TotalCases")+
  geom_line(color = "Yellow")+
  geom_point()

#Confidence Interval for mean population GrowthRate:
meanGR = mean(dataTrend1$GrowthRate)
sdGR = sd(dataTrend1$GrowthRate)

uppperlimitGR= meanGR+sdGR
lowerlimitGR = meanGR-sdGR

#Predictions based on: Current Growth Rate, MeanGrowth Rate, Upper GrowthRate, Lower GrowthRate
pred_val_Current <- (head(absolutenumbers, n = 1)*multiplier[,"GrowthRate"])+head(absolutenumbers, n = 1)
pred_val_MeanGR <- (head(absolutenumbers, n = 1)*meanGR)+head(absolutenumbers, n = 1)
pred_val_upperGR <- (head(absolutenumbers, n = 1)*uppperlimitGR)+head(absolutenumbers, n = 1)
pred_val_lowerGR <- (head(absolutenumbers, n = 1)*lowerlimitGR)+head(absolutenumbers, n = 1)

print("Current GR:")
print(multiplier[,"GrowthRate"])
print(pred_val_Current)

print("Mean GR:")
print(meanGR)
print(pred_val_MeanGR)

print("Upperlimit GR:")
print(uppperlimitGR)
print(pred_val_upperGR)

print("Lowerlimit GR:")
print(lowerlimitGR)
print(pred_val_lowerGR)

#Doubling rate:-
DRCurrent <- round(70/(meanGR*100),digits = 0)
DRMedian <- round(70/(median(dataTrend1$GrowthRate)*100),digits = 0)
latestDRCurrent <- round(tail(dataTrend1$doublinrRate,n=1),digits = 0)

```

```{r echo=FALSE,warning=FALSE}
#grid.arrange(p3,p3_log,p1,p2,p4,p5,p6,nrow = 4, ncol = 2)
print(p3)
print(p3_log)
print(p1)
print(p2)
print(p4)
print(p5)
print(p6)
print(p8)
print(p9)
print(p10)
print(p7)

```

### Results:
- Tomorrows predicted case number is between **`r pred_val_Current`** and **`r pred_val_MeanGR`** cases.
- Current doubling rate is **`r latestDRCurrent` days**
- Mean doubling rate, i.e number of cases doubles every **`r DRCurrent` days**.
- Median doubling rate, i.e number of cases doubles every **`r DRMedian` days**.
- Latest Observed Growth Rate **`r multiplier[,"GrowthRate"]*100`** percent.
- Mean Growth Rate: **`r meanGR*100`** percent.
- Median Growth Rate: **`r median(dataTrend1$GrowthRate)*100`** percent.
- Standard deviation: +/- **`r sdGR*100`**.
